\\\n **fabryka**

### 2025-08-07 fabryka_pci/article.Rmd
peer review completed (PCI archaeology)

---
title: '**fabryka**'
subtitle: A web application for easily analysing & exploring the fabric of archaeological
  assemblages
output:
  word_document: default
  pdf_document:
    keep_tex: true
header-includes:
- \usepackage{amsmath}
- \usepackage{lineno}
- \linenumbers
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{fabryka}
- \fancyhead[R]{R web application for fabric analysis}
- \fancyfoot[C]{\thepage}
- \usepackage{float}
- \floatplacement{figure}{H}
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
bibliography: references.bib
link-citations: true
---

@@

\begin{figure} [H]
\centering
  \includegraphics[width=4cm]{figure/hex_Fabryka4.png}
\label{}
\end{figure}

Marc THOMAS ^1\*^ (<https://orcid.org/0000-0002-8160-1910>) \smallbreak

@@

E-mail: [marcthomas1\@hotmail.fr](mailto:marcthomas1@hotmail.fr){.email}
\bigbreak

**Key words** : R shiny web application, fabric, site formation
processes, spatial analysis, orientation and dip \bigbreak

@@

During excavation, the orientation and dip of elongated remains (*i.e.* a > 2 * b and with a > 2 cm) is
measured on a representative set of artefacts or on the whole
assemblage. The orientation (= bearing) of the lowest point is measured
in the range [0; 360[ with North as the reference (*i.e.* orientation =
0°). The dip (= plunge) is measured in the range [0; 90] with horizontal
as a reference. The orientation [0; 360[ can be converted into axial
data measured in the range [0; 180[ by subtracting 180° from the
measurements in the range [180; 360].

@@

# Useful links and tips

-   use fabryka online: <https://marchaeologist.shinyapps.io/fabryka/> \smallbreak
-   source code: <https://github.com/marchaeologist/fabryka/tree/main> \smallbreak
-   Report bugs or any request: <https://github.com/marchaeologist/fabryka/issues> \smallbreak
-   github reprository doi of fabryka: <https://doi.org/10.5281/zenodo.15909666> \smallbreak
-   Zenodo repository of this article: <https://doi.org/10.5281/zenodo.16601423> \bigbreak

@@

The writing of the script was greatly facilitated by the work published
by Shannon McPherron (2018). It includes some functions published by the
author, with a few modifications. This work benefited from discussions
with Jean-Pierre Texier, Arnaud Lenoble, Pascal Bertran and Dominique
Todisco on issues related to site formation processes. The reference
data used to build the models of the classical method were collected by
P. Bertran throughout his career. Pascal Bertran provided us with all these
data in order to feed the application. The development of the
application also benefited from the help of Rémi Lemoy, François Baleux and
Victor Baleux on mathematical issues. A first version of the application
was tested by Emmanuel Discamps and Aurélien Royer. This project is
supported by the "Service Valorisarion de la Recherche" of the
University Toulouse Jean Jaurès. Finally, the constructive comments of
Pascal Bertran, recommender of this article, and, of Frédéric
Santos, Nicolas Frerebeau and Alfonso Benito-Calvo, reviewers of this
article, considerably improved the manuscript and the web application. I
would like to thank all these reasearchers for their advice. \bigbreak

@@

# Bibliography

The 'Bibliography' panel in the application contains the bibliography
cited in the application. This article contains the same bibliography
with some additional references. \bigbreak

>>>>

---
title: "\\\n **fabryka**"
subtitle: A web application for easily analysing & exploring the fabric of archaeological
  assemblages
output:
  pdf_document:
    keep_tex: true
  word_document: default
header-includes:
- \usepackage{titling}
- \pretitle{\begin{flushright}
- \includegraphics[width=2in,height=2in]{logo.png}\LARGE\\}
- \posttitle{\end{flushright}}
- \usepackage{amsmath}
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{fabryka}
- \fancyhead[R]{R web application for fabric analysis}
- \fancyfoot[C]{\thepage}
- \usepackage{float}
- \floatplacement{figure}{H}
editor_options:
  markdown:
    wrap: 72
  chunk_output_type: console
geometry: margin=2.54cm
bibliography: references.bib
link-citations: true
---

@@

**Marc Thomas** ^1\*^ (<https://orcid.org/0000-0002-8160-1910>) \smallbreak

@@

E-mail: [marcthomas1\@hotmail.fr](mailto:marcthomas1@hotmail.fr){.email}
\smallbreak

**Key words**: R shiny web application, fabric, site formation
processes, spatial analysis, orientation and dip \bigbreak

@@

During excavation, the orientation and dip of elongated remains (*i.e.*
a \> 2 \* b and with a \> 2 cm) is measured on a representative set of
artefacts or on the whole assemblage. The orientation (= bearing) of the
lowest point is measured in the range [0; 360[ with North as the
reference (*i.e.* orientation = 0°). The dip (= plunge) is measured in
the range [0; 90] with horizontal as a reference. The orientation [0;
360[ can be converted into axial data measured in the range [0; 180[ by
subtracting 180° from the measurements in the range [180; 360].

@@

# Data, scripts, code, and supplementary information availability

-   use fabryka online: <https://marchaeologist.shinyapps.io/fabryka/>
    \smallbreak
-   source code: <https://github.com/marchaeologist/fabryka/tree/main>
    \smallbreak
-   Report bugs or any request:
    <https://github.com/marchaeologist/fabryka/issues> \smallbreak
-   github reprository doi of fabryka:
    <https://doi.org/10.5281/zenodo.15909666> \smallbreak
-   Zenodo repository of this article:
    <https://doi.org/10.5281/zenodo.16758152> \bigbreak

@@

The writing of the script was greatly facilitated by the work published
by Shannon McPherron (2018). It includes some functions published by the
author, with a few modifications. This work benefited from discussions
with Jean-Pierre Texier, Arnaud Lenoble, Pascal Bertran, Thomas Perrin and Dominique
Todisco on issues related to site formation processes. The reference
data used to build the models of the classical method were collected by
P. Bertran throughout his career. Pascal Bertran provided us with all
these data in order to feed the application. The development of the
application also benefited from the help of Rémi Lemoy, François Baleux
and Victor Baleux on mathematical issues. A first version of the
application was tested by Emmanuel Discamps and Aurélien Royer. This
project is supported by the "Service Valorisarion de la Recherche" of
the University Toulouse Jean Jaurès. A preprint version of this article (v3, https://doi.org/10.5281/zenodo.16601423) has been peer-reviewed and recommended by PCI Archaeo (https://doi.org/10.24072/pci.archaeo.100611). Comments made by Pascal Bertran, recommender of this article, and, by Frédéric Santos, Nicolas Frerebeau and Alfonso Benito-Calvo, reviewers of this article, considerably improved the manuscript and the application. I would like to thank all these reasearchers for their advice. \bigbreak

# Funding

This research was funded by the Service Valorisation de la Recherche of Toulouse - Jean Jaurès University. \bigbreak

@@

# References

The 'Bibliography' panel in the application contains the references
cited in the application. This article contains the same bibliography
with some additional references. \bigbreak
